### 2003-07-15 docs/MSTK.tex
MSTK Mesh Toolkit by Rao V. Garimella (rao@lanl.gov), T-7, Los Alamos National Laboratory, Los Alamos, NM 87545

\documentclass[12pt]{article}

% TeX calculates margins from a point 1 inch from the top and 1 inch
% from the left of the page. So the total left margin for instance is
% the 'oddsidemargin' + 1 inch.
%
%\setlength{\textwidth}{4.5in}
%\setlength{\textheight}{7.0in}
%\setlength{\evensidemargin}{0.0in}
%\setlength{\oddsidemargin}{0.0in}
%\setlength{\topmargin}{0.5in}

\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{wrapfig}
\usepackage{algorithmic}
\usepackage[boxed]{algorithm}
\usepackage{rotating}
\usepackage{multicol}
%\usepackage{times}

%\displaybreak[0]       %% Break permitted but not encouraged

\begin{document}

\title{MSTK: Mesh Toolkit, v 1.0}

\author{Rao V. Garimella \\
  Mathematical Modeling and Analysis (T-7), Theoretical Division, \\
  Los Alamos National Laboratory, Los Alamos, NM, USA. \\ E-mail:
  rao@lanl.gov}

\maketitle

\thispagestyle{empty}

\newpage

\section{MSTK IO File Format}

\subsection{ASCII File Format}

\begin{verse}
\# \textit{This is a comment} \\
\# \textit{The string ``MSTK'' and File version number (1.0)} \\
\vspace{1ex}

MSTK Ver %
\vspace{1ex}

\# \textit{char *Mfilename - Geometric model file related to mesh} 
\vspace{1ex}

Mfilename
\vspace{1ex}

\# \textit{char *reptype - Type of representation} \\  
\# \textit{int NV, NE, NF, NR - Number of vertices, edges, face, regions} \\
\vspace{1ex}

RepType \hspace{0.5ex} NV \hspace{0.5ex} NE \hspace{0.5ex} NF \hspace{0.5ex} NR \\
\hspace{2ex}

\# \textit{\textbf{VERTEX INFO}} \\
\# \textit{Each record has} \\
\# \textit{double X,Y,Z Coordinates} \\
\# \textit{int Mdim - Topological type or dimension of model entity that} \\
\# \textit{\hspace{4.7em} the vertex is on} \\
\# \textit{int Mid - ID of model entity the vertex is on} \\
\# \textit{int Mdim and Mid can be -1 if model information is absent} 
\vspace{1ex}

X\_Coord \hspace{0.5ex} Y\_Coord \hspace{0.5ex} Z\_Coord \hspace{0.5ex} Mdim \hspace{0.5ex} Mid \\
X\_Coord \hspace{0.5ex} Y\_Coord \hspace{0.5ex} Z\_Coord \hspace{0.5ex} Mdim \hspace{0.5ex} Mid \\
. . . \\
\# \textit{Repeated NV times}
\vspace{2ex}

\# \textit{\textbf{EDGE INFO} - present only if NE $\ne$ 0} \\
\# \textit{Each edge record has} \\
\# \textit{int Vid\_1, Vid\_2  - IDs of first, second vertex of edge} \\
\# \textit{int Mdim, Mid}
\vspace{1ex}

Vid\_1 \hspace{0.5ex} Vid\_2 \hspace{0.5ex} Mdim \hspace{0.5ex} Mid \\
Vid\_1 \hspace{0.5ex} Vid\_2 \hspace{0.5ex} Mdim \hspace{0.5ex} Mid \\
. . . \\
\# \textit{Repeated NEdges times}
\vspace{2ex}

\newpage
\# \textit{\textbf{FACE INFO} - present only if NF $\ne$ 0} \\
\# \textit{char *FLtype - Dimension of lower order entity describing faces} \\
\# \textit{Values: Vertex, Edge (case insensitive), e.g. VeRteX or EDGE} \\
\vspace{1ex}

FLtype 
\vspace{1ex}

\# \textit{If face described by vertices, then each face record has} \\
\# \textit{int NFV - Number of face vertices} \\
\# \textit{int Vid\_1 - ID of first vertex of face} \\
\# \textit{int Vid\_2 - ID of second vertex of face} \\
\# . . . \\
\# \textit{int Vid\_1 - ID of NFV'th vertex of face} \\
\# \textit{int Mdim, Mid} 
\vspace{1ex}

NFV \hspace{0.5ex} Vid\_1 \hspace{0.5ex} Vid\_2 \hspace{0.5ex} ... \hspace{0.5ex} Vid\_NFV \hspace{0.5ex} Mdim \hspace{0.5ex} Mid \\
NFV \hspace{0.5ex} Vid\_1 \hspace{0.5ex} Vid\_2 \hspace{0.5ex} ... \hspace{0.5ex} Vid\_NFV \hspace{0.5ex} Mdim \hspace{0.5ex} Mid \\
. . . \\
\# \textit{Repeated NFaces times}
\vspace{2ex}\vspace{1ex}

\# \textit{If face described by edges, then each face record has} \\
\# \textit{int NFE - Number of face edges} \\
\# \textit{int $\pm$Eid\_1 - signed ID of first edge of face} \\
\# \textit{int $\pm$Eid\_2 - signed ID of second edge of face} \\
\# . . . \\
\# \textit{int $\pm$Eid\_NFE - signed ID of NFE'th edge of face} \\
\# \textit{int Mdim, Mid} \\
\# \\
\# \textit{if sign of edge is +, face uses edge in direction it was defined} \\
\# \textit{if sign of edge is -, face uses edge in opposite direction} \\
\vspace{1ex}

NFE \hspace{0.5ex} $\pm$Eid\_1 \hspace{0.5ex} $\pm$Eid\_2 \hspace{0.5ex} ... \hspace{0.5ex} $\pm$Eid\_NFE \hspace{0.5ex} Mdim Mid \\
NFE \hspace{0.5ex} $\pm$Eid\_1 \hspace{0.5ex} $\pm$Eid\_2 \hspace{0.5ex} ... \hspace{0.5ex} $\pm$Eid\_NFE \hspace{0.5ex} Mdim Mid \\
. . . \\
\# \textit{Repeated NFaces times}
\vspace{2ex}\vspace{1ex}

\# \textit{\textbf{REGION INFO} - present only if NR $\ne$ 0} \\
\# \textit{char *RLtype - Dimension of Lower order entity describing region} \\
\# \textit{Values: Vertex, Face (case insensitive), e.g. VERtex or faCE}
\vspace{1ex}

RLtype
\vspace{1ex}

\# \textit{if region described by vertices, then each region record has} \\
\# \textit{int NRV - Number of region vertices} \\
\# \textit{int Vid\_1 - ID of first vertex of region} \\
\# \textit{int Vid\_2 - ID of second vertex of region} \\
\# . . . \\
\# \textit{int Vid\_NFE - ID of NRV'th vertex of region} \\
\# \textit{int Mid, (NOTE: Mdim is not specified, since it has to be 3)}
\vspace{1ex}

NRV \hspace{0.5ex} Vid\_1 \hspace{0.5ex} Vid\_2 \hspace{0.5ex} ... \hspace{0.5ex} Vid\_NRV \hspace{0.5ex} Mid \\
NRV \hspace{0.5ex} Vid\_1 \hspace{0.5ex} Vid\_2 \hspace{0.5ex} ... \hspace{0.5ex} Vid\_NRV \hspace{0.5ex} Mid \\
. . . \\
\# \textit{Repeat NR times} 
\vspace{3ex}

\# \textit{if region described by faces, then each region record has} \\
\# \textit{int NRF - Number of region faces} \\
\# \textit{int Fid\_1 - signed ID of first face of region} \\
\# \textit{int Fid\_2 - signed ID of second face of region} \\
\# . . . \\
\# \textit{int Fid\_NRF - signed ID of NRF'th face of region} \\
\# \textit{int Mdim, Mid} \\
\#  \\
\# \textit{if sign of face is +, face normal points out of region} \\
\# \textit{if sign of edge is -, face normal points into region}
\vspace{1ex}

NRF \hspace{0.5ex} $\pm$Fid\_1 \hspace{0.5ex} $\pm$Fid\_2 \hspace{0.5ex} ... \hspace{0.5ex} $\pm$Fid\_NRF \hspace{0.5ex} Mid \\
NRF \hspace{0.5ex} $\pm$Fid\_1 \hspace{0.5ex} $\pm$Fid\_2 \hspace{0.5ex} ... \hspace{0.5ex} $\pm$Fid\_NRF \hspace{0.5ex} Mid \\
. . . \\
\# \textit{Repeated NR times} 
\vspace{2ex}

\# \textit{\textbf{VERTEX ATTRIBUTES}} \\
\# \textit{int NVA - Number of Vertex attributes} \\
\# \\
\# \textit{char *VA\_name\_1 - Name of first vertex attribute} \\
\# \textit{int VA\_type\_1 - Type of first vertex attribute} \\
\# \textit{int VA\_dim\_1 - Dimension of first vertex attribute} \\
\# \\
\# \textit{char *VA\_name\_2 - Name of second vertex attribute} \\
\# \textit{int VA\_type\_2 - Type of second vertex attribute} \\
\# \textit{int VA\_dim\_2 - Dimension of first vertex attribute} \\
\# \\
\# . . .\\
\# \\
\# \textit{char *VA\_name\_NVA - Name of NVA'th vertex attribute} \\
\# \textit{int VA\_type\_NVA - Type of NVA'th vertex attribute} \\
\# \textit{int VA\_dim\_NVA - Dimension of NVA'th vertex attribute} \\
\# \\
\# \textit{VA\_type can be 1 (int), 2 (double), 3 (string)} \\
\# \textit{VA\_dim = 1 for scalars, VA\_dim = length of vector for vector} \\
\# \textit{VA\_dim can only be 1 when VA\_type is string}
\vspace{1ex}

NVA \hspace{0.5ex} \\
VA\_name\_1 \hspace{0.5ex} VA\_type\_1 \hspace{0.5ex} VA\_dim\_1 \\ 
VA\_name\_2 \hspace{0.5ex} VA\_type\_2 \hspace{0.5ex} VA\_dim\_2 \\
. \\
VA\_dim\_NVA \hspace{0.5ex} VA\_type\_NVA \hspace{0.5ex} VA\_dim\_2  
\vspace{1ex}

\# \textit{For each vertex attribute record, list of attribute values} \\
\# \textit{E.G., there are 3 attributes for each vertex:} \\
\# \textit{a scalar int, a vector of 3 doubles and a string}

VA\_int \hspace{0.5ex} VA\_double\_1 \hspace{0.5ex} VA\_double\_2 \hspace{0.5ex} VA\_double\_3 \hspace{0.5ex} VA\_string \\
VA\_int \hspace{0.5ex} VA\_double\_1 \hspace{0.5ex} VA\_double\_2 \hspace{0.5ex} VA\_double\_3 \hspace{0.5ex} VA\_string \\
. \\
. \\
. \\
\# \textit{Repeated NV times}
\vspace{2ex}

\newpage
\# \textit{\textbf{EDGE ATTRIBUTES}} \\
\# \textit{Similar to Vertex attribute description}
\vspace{1ex}

NEA \hspace{0.5ex} \\
EA\_name\_1 \hspace{0.5ex} EA\_type\_1 \hspace{0.5ex} EA\_dim\_1 \\ 
EA\_name\_2 \hspace{0.5ex} EA\_type\_2 \hspace{0.5ex} EA\_dim\_2 \\
. \\
EA\_dim\_NEA \hspace{0.5ex} EA\_type\_NEA \hspace{0.5ex} EA\_dim\_2  
\vspace{1ex}

\# \textit{For each edge attribute record, list of attribute values} \\
\# \textit{E.G., a scalar int, a scalar double and a string}
\vspace{1ex}

EA\_int \hspace{0.5ex} EA\_double \hspace{0.5ex} EA\_string \\
EA\_int \hspace{0.5ex} EA\_double \hspace{0.5ex} EA\_string \\
. . . \\
\# \textit{Repeated NE times}
\vspace{2ex}

\# \textit{\textbf{FACE ATTRIBUTES}} \\
\# \textit{Similar to Vertex attribute description}
\vspace{1ex}

NFA \hspace{0.5ex} \\
FA\_name\_1 \hspace{0.5ex} FA\_type\_1 \hspace{0.5ex} FA\_dim\_1 \\ 
FA\_name\_2 \hspace{0.5ex} FA\_type\_2 \hspace{0.5ex} FA\_dim\_2 \\
. \\
FA\_dim\_NFA \hspace{0.5ex} FA\_type\_NEA \hspace{0.5ex} FA\_dim\_2  
\vspace{1ex}

\# \textit{For each face attribute record, list of attribute values} \\
\# \textit{E.G., a vector of 2 doubles and a string}
\vspace{1ex}

FA\_double\_1 \hspace{0.5ex} FA\_double\_2 \\
FA\_double\_1 \hspace{0.5ex} FA\_double\_2 \\
. . . \\
\# \textit{Repeated NF times}
\vspace{2ex}

\newpage
\# \textit{\textbf{REGION ATTRIBUTES}} \\
\# \textit{Similar to Vertex attribute description}
\vspace{1ex}

NRA \hspace{0.5ex} \\
RA\_name\_1 \hspace{0.5ex} RA\_type\_1 \hspace{0.5ex} RA\_dim\_1 \\ 
RA\_name\_2 \hspace{0.5ex} RA\_type\_2 \hspace{0.5ex} RA\_dim\_2 \\
. \\
RA\_dim\_NRA \hspace{0.5ex} RA\_type\_NRA \hspace{0.5ex} RA\_dim\_NRA 
\vspace{1ex}

\# \textit{For each Region attribute record, list of attribute values} \\
\# \textit{E.G., a vector of 3 ints}
\vspace{1ex}

RA\_int\_1 \hspace{0.5ex} RA\_int\_2 \hspace{0.5ex} RA\_int\_3 \\
RA\_int\_1 \hspace{0.5ex} RA\_int\_2 \hspace{0.5ex} RA\_int\_3 \\
. . .\\
\# \textit{Repeated NR times}
\vspace{2ex}

\end{verse}

\end{document}